$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-09-07 paper/paper.tex
More general type for the example derivation

The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous processes (\autoref{fig:example-process}).
The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
Thanks for linear and graded types, the processes below have their interactions safely restricted: the usage annotation $(0 \comma 1)$ on channel $c$ (no input, output once) restricts what both $\func{send}$ subprocesses can do; and $(2 \comma 0)$ (input twice, no output) restricts what $\func{send}$ can do.
Dually, the remaining annotations $(1 \comma 0)$, $(1 \comma 0)$ and $(0 \comma 2)$ are given to the subprocess $\func{sync}$.
We use these types to give a suitable typing derivation in \autoref{example-derivation}.

@@

We present an example with four parallel processes: two processes that send data over a channel; one that receives data over a channel twice; and a fourth process that sticks the previous processes together and synchronizes them.

We start by defining the process doing the synchronization.
This process sequentially receives data on both input channels and then sends on the output channel both pieces of received data.
For simplicity, when we type this process we instantiate it to variable references and usage annotations.
We expect both input channels to be graded and capable of performing input once.
The output channel must be capable of performing output twice.
Notice that the well-typedness proof has no choice but to blindly follow the syntax.

\begin{flalign*}
& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (1 \comma 0) \comma (1 \comma 0) \comma (0 \comma 2)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (0 \comma 0) \comma (0 \comma 0) \comma (0 \comma 0)} & \\
& \func{\vdash sync} && = \\
& && \recv{\suc \suc \constr{0}}{} \\
& && \recv{\suc \suc \constr{0}}{} \\
& && \send{\suc \suc \constr{0}}{\suc \constr{0}} \\
& && \send{\suc \suc \constr{0}}{\constr{0}} \PO \\
\end{flalign*}

Next, we define the processes doing the input and output.
This time too, we instantiate them to concrete variable references for typing.

\begin{flalign*}
& \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (2 , 2)}{\func{recv} \; \constr{0}}{\Gamma \comma (0 \comma 2)} & \\
& \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
& \\
& \func{\vdash send} \; \types{\gamma \comma \channel{t}{\lz} \comma t}{\Gamma \comma (1 , 1) \comma \lz}{\func{send} \; (\suc \constr{0}) \; \constr{0}}{\Gamma \comma (1 \comma 0)\comma \lz} & \\
& \func{\vdash send} \; = \; \send{(\suc \constr{0})}{\constr{0}} \PO
\end{flalign*}

@@

\begin{flalign*}
& \func{\vdash example} && : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}} & \\
& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
& && \comp{}{ \new{} \; \channel{\unit}{\func{\omega}} \; \_ \; \_ } \; ( \func{\vdash recv} \\
& && \comp{}{ \func{\vdash sync} }))) \\
\end{flalign*}

>>>>

The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous processes (\autoref{fig:example-process}).
The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
Thanks for linear and graded types, the processes below have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
We use these types to give a suitable typing derivation in \autoref{example-derivation}.

@@

We provide an example typing derivation for the processes defined in \autoref{example-process}.
For the sake of simplicity, we instantiate these processes with concrete variable references before typing them.

\begin{flalign*}
  & \func{\vdash send} \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}}{\Gamma \comma (l \comma \suc r)}{\func{send} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
  & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO \\
  & \\
  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
\end{flalign*}

We expect both input channels to be graded and capable of performing input once.
The output channel must be capable of performing output twice.
Notice that the well-typedness proof has no choice but to blindly follow the syntax.

\begin{flalign*}
& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
& \func{\vdash sync} && = \\
& && \recv{\suc \suc \constr{0}}{} \\
& && \recv{\suc \suc \constr{0}}{} \\
& && \send{\suc \suc \constr{0}}{\suc \constr{0}} \\
& && \send{\suc \suc \constr{0}}{\constr{0}} \PO \\
\end{flalign*}

@@

\begin{flalign*}
& \func{\vdash example} && : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}} & \\
& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; ( \func{\vdash send} \\
& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash send} \\
& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash recv} \\
& && \comp{}{ \func{\vdash sync} }))) \\
\end{flalign*}
